Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2023-12-20 doc/manuscript/translocation_bioarxiv_SI.tex

\templatetype{pnassupportinginfo}
\usepackage{longtable}

\title{Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease}
\author{Roland A. Knapp, Mark Q. Wilber, Allison Q. Byrne, Maxwell B. Joseph, Thomas C. Smith, Andrew P. Rothstein, Robert L. Grasso, Erica Bree Rosenblum}
\correspondingauthor{Corresponding Author name. Roland A. Knapp \\ E-mail: roland.knapp@ucsb.edu}

@@

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext
\hypertarget{frog-evolution-in-response-to-bd-2}{%
\section{Frog evolution in response to
Bd}\label{frog-evolution-in-response-to-bd-2}}

@@

We compared frog genomes sampled in naive versus recovering populations.
Bd-exposure histories of these populations are based on a decade or more
of data from visual encounter surveys and Bd surveillance using skin
swabbing \citep[e.g.,][]{knapp2016}. Comparing populations with
different infection histories allowed larger sample sizes and
replication across the landscape. The alternative approach of comparing
samples from the same populations before and after Bd exposure isn't
feasible in this system because Bd arrived in most MYL frog populations
decades ago and population persistence/recovery is rare and
unpredictable. As a result, samples from recovering populations
collected from before and after Bd exposure are not available and are
unlikely to be available in the future.

@@

In the Results section of the main paper, we describe the stringent set
of outlier SNPs (identified using a Bonferroni-corrected p-value of
0.01). The liberal set, identified using a Bonferroni-corrected p-value
of 0.05, included 38 outliers (35 SNPs and 3 INDELS) from 30 distinct
genes across 16 contigs. Our analysis found no significantly
over-represented GO terms in this set of 30 genes, or in the set of 38
outlier regions identified with the splined window analysis.

In the comparison of individual outlier variants between naive and
recovering populations, we detected consistent allele frequency changes
across all genetic clusters for only 1 of 7 of the associated genes
(RIN3; Fig. 1). However, for the other 6
genes we still found a statistically significant signal of selection in
2 of the 3 genetic clusters (containing populations 6--9;
Fig. 1, Figure~\ref{fig-selectionresults}
A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe parallel
evolutionary changes following Bd exposure and are therefore interesting
candidates for future studies.

\hypertarget{frog-population-recovery-2}{%
\section{Frog population
recovery}\label{frog-population-recovery-2}}

\hypertarget{methods-1}{%
\subsection{Methods}\label{methods-1}}

\hypertarget{population-viability-modeling-1}{%
\section{Population viability
modeling}\label{population-viability-modeling-1}}

\hypertarget{methods-2}{%
\subsection{Methods}\label{methods-2}}

We extracted yearly survival probabilities for translocated adults
\(\sigma_{A_T}\) and naturally recruited adults \(\sigma_{A_R}\) from
the CMR model used to estimate population-level parameters of
translocated populations (see ``Materials and Methods - Estimation of
frog survival and abundance'' in main paper). Although we observed
yearly variability in adult survival within a population, the magnitude
of this variability was small compared to among-population variability
(Fig. 3). Thus, we did not include
yearly within-population variability in adult survival in this analysis.
However, within a population there was substantial yearly variability in
the successful recruitment of adults, greater than what we would expect
from Poisson variability around a mean value. Therefore, we allowed for
yearly variability in the probability of successful recruitment
\(\omega\) (additional details provided in ``Estimating model
parameters'' below).

The baseline parameter values for the model and how they were estimated
are given in Table~\ref{tbl-param_values}. Parameters \(\sigma_{A_T}\)
and \(\sigma_{A_R}\) were extracted directly from our CMR models (see
``Materials and Methods - Estimation of frog survival and abundance'' in
main paper). For populations where we had a sufficient number of
PIT-tagged, naturally-recruited adults, we observed that
\(\sigma_{A_T}\) and \(\sigma_{A_R}\) could be notably different, with
\(\sigma_{A_R} > \sigma_{A_T}\) (Figure~\ref{fig-compare_surv_probs}).
For populations lacking sufficient numbers of naturally-recruited
adults, we were unable to directly estimate \(\sigma_{A_R}\), and
instead set \(\sigma_{A_R} = \sigma_{A_T}\).

We performed four analyses on our model. First, we considered a
deterministic version of our model with no yearly heterogeneity in
probability of successful recruitment \(\omega\), and calculated the
predicted long-run growth rate \(\lambda\) of a population for different
values of \(\sigma_{A_R}\) and \(\omega\). We then fixed
\(\omega = 0.3\) and calculated the predicted growth rate of our 12
populations.

Second, we performed a local elasticity analysis on \(\lambda\) with
respect to parameters \(\omega\), \(\sigma_{A_R}\), and \(F\) to
determine how small changes in these parameters could influence the
long-run deterministic growth rate of populations
(Figure~\ref{fig-viability-supp}).

Third, we defined a version of the model with demographic and
environmental stochasticity, where environmental stochasticity was
represented by among-year variability in \(\omega\). We used this model
to simulate a one-time introduction of 40 translocated adult frogs. We
ran this simulation 1000 times for each population and computed the
probability of a population becoming extinct after 50 years given the
observed parameter values and environmental stochasticity in \(\omega\).
We varied the mean recruitment probability \(\omega\) from 0.15 and 0.4
and drew values of \(\omega\) each year from a beta distribution with a
dispersion parameter of \(\phi = 2\) (when \(\omega = 0.5\) and
\(\phi = 2\) the beta distribution is uniform between 0 and 1).

Finally, we assessed whether our stochastic model could reproduce
observed trajectories of population recovery. We focused on population
70550 because this was our longest CMR time series for a translocated
population and because this population shows evidence of substantial
post-translocation increases in adult abundance associated with
population establishment and recovery. We simulated our model for 16
years, repeating the simulation 50,000 times. For each run and each
year, we drew \(\omega\) from a uniform distribution between 0 and 1 (or
equivalently a beta distribution with mean 0.5 and \(\phi = 2\)). Using
Approximate Bayesian Computing and rejection sampling
\citep{kosmala2016}, we identified the top 2\% of trajectories (i.e.,
1000 trajectories) that minimized the sum of squared errors between the
observed and predicted data. The \(\omega\) values associated with these
``best'' trajectories represented an approximate posterior distribution
\citep{beaumont2010}. Using these best fit trajectories, we assessed
whether our model could qualitatively describe the patterns of recovery
in the observed data for population 70550.

@@

\caption{\label{fig-synteny-plot}Synteny plot showing conserved gene
order in \emph{Xenopus tropicalis} and \emph{Rana mucosa} for the
outlier region containing MHC Class I Classical and MHC Class III gene
regions. The plot was created with SimpleSynteny \citep{veltri2016}
using \emph{Xenopus tropicalis} Chromosome 8 (NC\_030684.2, genbank
accession GCA\_000004195.4) and \emph{Rana mucosa} Contig19. Asterices
indicate the location of SNP outliers in the TCF19 and VARS1 genes. Gap
sizes for each contig representation are labeled.}

@@

@@

\caption{\label{fig-yosemap}Map showing the locations of translocated
and donor MYL frog populations in Yosemite National Park (park boundary
indicated by black polygon). Symbol shapes indicate the donor population
used for each translocation site. To obscure the exact locations of
populations, random noise was added to all point coordinates. Inset map
shows the location of Yosemite within California. In both maps,
elevation is indicated by the colored hillshade layer (dark green =
lowest elevation, white = highest elevation).}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/mcmc_areas_m2b.png}

@@

\caption{\label{fig-transsurvival-postdens}For the best model describing
the average 1-year survival of frogs in the first translocation to a
site as a predictor of survival of frogs in all subsequent
translocations to that site, estimated posterior density curves and
shaded 95\% uncertainty intervals for the intercept and single predictor
variable. In the Bayesian framework in which the model was developed, a
variable is considered an important predictor if the associated
uncertainty interval does not overlap zero (indicated by the dashed red
line).}

@@

@@

\caption{\label{fig-bdload-beforeafter}Evidence that frogs translocated
from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. For frogs translocated to
each of the 12 recipient sites, Bd loads for the period immediately
prior to translocation versus during the 1-year period after
translocation. Bd loads are expressed as the number of ITS1 copies per
skin swab, as estimated by qPCR of the Bd ITS1 region. Box plots show
medians, first and third quartiles, largest and smallest values within
1.5x interquartile range, and values outside the 1.5x interquartile
range. Loads indicative of severe disease are \textgreater{} 5.8 ITS
copies (on a log\textsubscript{10} scale).}

\caption{\label{fig-survival-postdens}Evidence that Bd load is not an
important predictor of post-translocation frog survival. For the best
model identifying predictors of 1-year frog survival following
translocation, estimated posterior density curves and shaded 95\%
uncertainty intervals for the intercept and all predictor variables. In
the Bayesian framework in which the model was developed, variables are
considered important predictors if the associated uncertainty interval
does not overlap zero (indicated by the dashed red line). Predictor
variables shown on the y-axis are defined as follows: length = frog
size, snow\_t = winter severity in the year of translocation (measured
on April 1), snow\_t1 = winter severity in the year following
translocation (measured on April 1), day = day of year on which a
translocation was conducted, bd\_load = Bd load, elevation = site
elevation, order\_first = within-site translocation order, sex\_male =
frog sex, donor\_70567 and donor\_72996 = donor population.}

\caption{\label{fig-compare_surv_probs}Evidence that frog
resistance/tolerance in non-natal habitats is maintained across
generations. Comparison of the average yearly adult survival
probabilities for adults translocated to each of 3 sites versus adults
that were naturally recruited at each site (as a result of reproduction
by translocated frogs). In contrast to
Fig. 3, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohorts. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. All estimates were derived using
the mrmr package.}

>>>>

\templatetype{pnassupportinginfo}
\usepackage{longtable}
\usepackage{blkarray}

\title{Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease}
\author{Roland A. Knapp, Mark Q. Wilber, Allison Q. Byrne, Maxwell B. Joseph, Thomas C. Smith, Andrew P. Rothstein, Robert L. Grasso, Erica Bree Rosenblum}
\correspondingauthor{Corresponding Author name. Roland A. Knapp \\ E-mail: roland.knapp@ucsb.edu}

@@

%% Adds the main heading for the text. Comment out this line if you do not have any supporting information text.
\SItext
\hypertarget{frog-population-recovery-2}{%
\section{Frog population
recovery}\label{frog-population-recovery-2}}

\hypertarget{laboratory-methods}{%
\subsection{Laboratory Methods}\label{laboratory-methods}}

\hypertarget{cmr-model-structure}{%
\subsection{CMR model structure}\label{cmr-model-structure}}

We estimated survival and recruitment for each site using open
population CMR models based on \citep{joseph2018}. For each individual
\(i=1, ..., M\) on each survey \(j=1, ..., n_j\): \(o_{i, j} = 1\) if
the individual was not detected, and \(o_{i, j}=2\) if the individual
was detected. Capture histories of \(M\) individuals are modeled, though
only \(N_s\) individuals were captured.

This parameter expanded data augmentation allows us to capture the
possibility that undetected individuals may have recruited into the
adult population \citep{royle2012}. Here, \(M\) was chosen to be three
times the number of observed individuals (\(3N_s\)) to be considerably
greater than our prior guess of \(N_s\).

We denote the true state of individual \(i\) as \(u_{i, t}\) for primary
period \(t = 1,..., n_t\). The four states that we consider are:
\(u_{i, t} = 1\) for individuals that have not recruited,
\(u_{i, t} =2\) for live adults, and \(u_{i, t} = 3\) for dead adults.
Each survey \(j=1, ..., n_j\) occurs in one of the \(n_t\) primary
periods, and we denote the primary period in which survey \(j\) takes
place as \(t_j\). Each primary period \(t\) occurs within one year, but
within a year there can be multiple primary periods. We set the year
containing the first primary period to \(y_{t = 1} = 1\), and generally
\(y_t\) represents the year containing primary period \(t\). Years
increment by one until the final year of the mark recapture efforts,
which we denote \(n_y\): \(y \in \{1, 2, ..., n_y\}\). We assume that
within a primary period, the state of each individual does not change
(i.e., individuals do not recruit into the adult population, gain or
lose Bd infection, or die). This assumption is justified by the short
time intervals between surveys within primary periods, in cases where
primary periods contain multiple secondary periods.

Live individuals are detected with probability \(p_j\), which is modeled
as:

\[p_j = \text{logit}^{-1}(X_j^{(p)} \beta^{(p)}),\]

where \(X_j^{(p)}\) is a known row vector and \(\beta^{(p)}\) an unknown
parameter vector. Not recruited and dead individuals are never captured.
We bundle these assumptions about the observation probabilities for
survey \(j\) into an emission matrix \(\Omega_j\):

\[
\Omega_j =
\begin{blockarray}{ccc}
  \text{Not detected} & \text{Detected} \\
\begin{block}{(cc)c}
  1 & 0 & \text{Not recruited} \\
  1 - p_j & p_j & \text{Alive} \\
  1 & 0 & \text{Dead} \\
\end{block}
\end{blockarray}
\]

The state transition matrix \(\Psi_{t, i}\) contains the probabilities
of individual \(i\) transitioning from state \(u_{i, t}\) (rows) to
\(u_{i, t+1}\) (columns) between primary period \(t\) and \(t+1\). For
unintroduced individuals which naturally recruit, this matrix is given
by:

\[
\Psi_{t, i} =
\begin{blockarray}{cccc}
  \text{Not recruited} & \text{Alive} & \text{Dead} \\
\begin{block}{(ccc)c}
  1 - \lambda_t & \lambda_t & 0 & \text{Not recruited} \\
  0 & \phi_t & 1 - \phi_t & \text{Alive} \\
  0 & 1 & 1 & \text{Dead} \\
\end{block}
\end{blockarray},
\] where \(\lambda_t\) is the probability of recruiting in time \(t\)
and \(\phi_t\) is the probability of survival in time \(t\).

For introduced individuals which have deterministic recruitment (they
recruit when introduced), the state transition matrix is given by:

\[
\Psi_{t, i} =
\begin{blockarray}{cccc}
  \text{Not recruited} & \text{Alive} & \text{Dead} \\
\begin{block}{(ccc)c}
  1 - I_{t, i} & I_{t, i} & 0 & \text{Not recruited} \\
  0 & \phi_t & 1 - \phi_t & \text{Alive} \\
  0 & 1 & 1 & \text{Dead} \\
\end{block}
\end{blockarray},
\] where \(I_{t, i}\) is a known indicator function for whether
individual \(i\) was introduced in primary period \(t\).

We allow recruitment probabilities to vary in time via random effects,
such that:

\[\lambda_t = \text{logit}^{-1}(\alpha^{(\lambda)} + \epsilon^{(\lambda)}_t),\]

where \(\alpha^{(\lambda)}\) is an intercept parameter and
\(\epsilon^{(\lambda)}_t\) is an adjustment for time \(t\).

Survival probabilities also vary in time, and as a function of known
covariates:

\[\phi_t = \text{logit}^{-1}(X^{(\phi)}_t \beta^{(\phi)} + \epsilon^{(\phi)}_t),\]

where \(X^{(\phi)}_t\) is a row vector of known covariates,
\(\beta^{(\phi)}\) is a column vector of unknown coefficients, and
\(\epsilon^{(\phi)}_t\) is an adjustment for time \(t\).

To complete the specification of the Bayesian model, we need to specify
priors for all unknown parameters. The recruitment parameter priors were
specified as follows:

\[\alpha^{(\lambda)} \sim N(0,1),\]
\[\sigma^{(\lambda)} \sim N_+(0,1),\]
\[\epsilon^{(\lambda)}_t \sim N(0, \sigma^{(\lambda)}),\] for periods
\(t=1, ..., T\). Here \(N\) represents the normal distribution and
\(N_+\) the half normal distribution with positive support.

Survival parameter priors were specified similarly as:

\[\beta^{(\phi)} \sim N(0, 1),\] \[\sigma^{(\phi)} \sim N_+(0, 1),\]
\[\epsilon^{(\phi)}_t \sim N(0, \sigma^{(\phi)}),\] for time
\(t=1, ..., T\).

The detection model coefficient vector also received a standard normal
prior \(\beta^{(p)} \sim N(0, 1)\).

We computed the likelihood of each individual capture history using the
forward algorithm, and we estimated the latent states using the
forward-backward algorithm \citep{zucchini2009, joseph2018}.

All of the code to specify and fit the model in Stan is available in the
open source mrmr package \citep{joseph2019}.

The joint distribution of the resulting model can be written as follows:

\begin{multline*}
[\alpha^{(\lambda)}, \sigma^{(\lambda)}, \epsilon^{(\lambda)}_{1:T}, \beta^{(\phi)}, \sigma^{(\phi)}, \epsilon^{(\phi)}_{1:T}, \beta^{(p)} \mid \pmb Y] \propto \prod_{i=1}^M [Y_i \mid \alpha^{(\lambda)}, \epsilon^{(\lambda)}_{1:T}, \beta^{(\phi)}, \epsilon^{(\phi)}_{1:T}, \beta^{(p)}] \times \\
\prod_{t=1}^T [\epsilon_t^{(\lambda)} \mid \sigma^{(\lambda)}] [\epsilon_t^{(\phi])} \mid \sigma^{(\phi)}] [\sigma^{(\lambda)}] [\sigma^{(\phi)}] [\alpha^{(\lambda)}] [\beta^{(\phi)}] [\beta^{(p)}],
\end{multline*}

where \(\pmb Y\) is an \(M \times T\) detection matrix, and \(Y_i\) the
capture history of individual \(i\).

\hypertarget{among-site-survival-modeling}{%
\subsection{Among-site survival
modeling}\label{among-site-survival-modeling}}

We used Bayesian generalized linear mixed models to investigate
predictors of survival among sites. The response \(y_i\) is binary,
representing a point estimate of whether individual \(i\) survived in
the year following translocation. We generated these point estimates by
rounding the posterior median of 1-year post-introduction survival for
each individual, and modeled the data using a Bernoulli distribution:

\[y_i \sim \text{Bernoulli}(p_i),\]

where \(p_i\) is the probability of survival.

We modeled variation in probabilities as follows:

\[\text{logit}(p_i) = \alpha + X_i \pmb \beta + \pmb \nu_{g[1:N]},\]

where \(\alpha\) is an intercept, \(X_i\) is a length \(K\) row vector
of predictors, \(\pmb \beta\) is a column vector of predictor effects,
and \(\pmb \nu\) a vector of group level random effects. Here \(g[i]\)
refers to the group \(g\) containing individual \(i\), and we estimate
an adjustment for each of the \(G\) groups (\(\nu_1, ..., \nu_G\)).

These models were fit using the \texttt{stan\_glmer()} function in the
rstanarm package, with default priors described below
\citep{rstanarm2022}. These priors are vague, but include data dependent
scaling as follows to account for different input variable scales.
However, as we standardized all predictor variables similarly to have
equal variance (by centering and dividing by twice the sample standard
deviation), the resulting priors are identical. Specifically, we have:

\[\alpha \sim \text{Normal}(0, 2.5),\] and

\[\beta_k \sim \text{Normal}(0, 5),\]

for \(k=1, ..., K\) where \(K\) is the number of predictor variables.

The default prior for group level adjustments \(\nu_1, ..., \nu_G\) in
rstanarm is a zero-mean Gaussian, where the covariance matrix is
constructed from a correlation matrix with an LKJ prior, and a vector of
variance parameters -- the decomposition of variance prior with unit
regularization, concentration, shape, and scale parameters
\citep{lewandowski2009}.

We drew posterior samples using Dynamic Hamiltonian Monte Carlo in Stan,
with four parallel chains, each run for 10,000 iterations, discarding
the first half of each chain as warm-up draws \citep{rstanarm2022}. We
used Rhat statistics and trace plots to verify convergence. We
considered models with different subsets of fixed and random effects,
and used approximate leave-one-out cross validation to identify the best
model \citep{vehtari2016}.

Finally, while it would theoretically be possible to estimate survival
covariate effects in a joint CMR model that integrates capture histories
across all sites, this was impractical due to computational requirements
of the CMR models (namely, run time and memory). In modeling survival
estimates from site-specific CMR analyses, we are in effect conducting
an among-site meta-analysis to better understand how site, cohort, and
individual level characteristics explain variation in survival after
translocation.

\hypertarget{population-viability-modeling-1}{%
\subsection{Population viability
modeling}\label{population-viability-modeling-1}}

We computed yearly survival probabilities for translocated adults
\(\sigma_{A_T}\) and naturally recruited adults \(\sigma_{A_R}\) from
the posterior distribution of individual state trajectories derived from
CMR models. Although we observed yearly variability in adult survival
within a population, the magnitude of this variability was small
compared to among-population variability
(Fig. 3). Thus, we did not include
yearly within-population variability in adult survival in this analysis.
However, within a population there was substantial yearly variability in
the successful recruitment of adults, greater than what we would expect
from Poisson variability around a mean value. Therefore, we allowed for
yearly variability in the probability of year 1 juvenile survival and
recruitment (additional details provided in \textbf{Estimating model
parameters} below). We also could have included environmental
stochasticity in year 2 juvenile survival and recruitment
\(\sigma_{J_2}\), but our elasticity analysis showed that this parameter
had little effect on host growth rate relative to \(\sigma_{J_1}\)
(Figure~\ref{fig-viability-supp}).

The baseline parameter values for the model and how they were estimated
are given in Table~\ref{tbl-param_values}. Parameters \(\sigma_{A_T}\)
and \(\sigma_{A_R}\) were extracted directly from our CMR models (see
\textbf{Materials and Methods - Estimation of frog survival and
abundance} and \textbf{Supporting Information - Frog population recovery
- Estimation of frog survival and abundance} for details). For
populations where we had a sufficient number of PIT-tagged,
naturally-recruited adults, we observed that \(\sigma_{A_T}\) and
\(\sigma_{A_R}\) could be notably different, with
\(\sigma_{A_R} > \sigma_{A_T}\) (Figure~\ref{fig-compare_surv_probs}). For populations lacking sufficient numbers of naturally-recruited
adults, we were unable to directly estimate \(\sigma_{A_R}\), and
instead set \(\sigma_{A_R} = \sigma_{A_T}\).

To estimate the \(\sigma_{A_R}\), we used the posterior distribution of
predicted true states for naturally-recruited individuals (1=not
recruited, 2=alive, 3=dead, as described in \emph{CMR model structure}),
then calculated the posterior probability of individuals surviving
between consecutive primary periods, conditional on being alive in the
first primary period (e.g., given a value of 2 (alive) in the first
primary period, how often was the value still 2 (alive) in the next
primary period compared to 3 (dead)?) . This yielded posterior
distributions for survival probabilities between primary periods.
However, because the time interval between primary periods differed, the
survival probabilities between different consecutive primary periods
were not directly comparable. To address this, we converted the survival
probabilities between each consecutive primary period to per day death
rates, propagating the uncertainty from the posterior distributions. We
then took a weighted average of these death rates, weighted by the time
interval between primary periods, to get the average per day death rate
over the entire CMR survey. We converted this per day death rate \(d\)
back to a yearly survival probability using
\(\exp(-d \times 365 \text{ days})\). We used the same procedure for
\(\sigma_{A_T}\) such that our estimates of average yearly survival
probability were comparable between \(\sigma_{A_R}\) and
\(\sigma_{A_T}\).

We performed four analyses on our model. First, we considered a
deterministic version of our model with no yearly heterogeneity in year
1 juvenile survival and recruitment probability \(\sigma_{J_1}\), and
calculated the predicted long-run growth rate \(\lambda\) of a
population for different values of \(\sigma_{A_R}\) and
\(\sigma_{J_1}\). We then fixed \(\sigma_{J_1} = 0.09\) and calculated
the predicted growth rate of our 12 populations.

Second, we performed a local elasticity analysis on \(\lambda\) with
respect to parameters \(\sigma_{J_1}\), \(\sigma_{J_2}\),
\(\sigma_{A_R}\), and \(F\) to determine how small changes in these
parameters could influence the long-run deterministic growth rate of
populations (Figure~\ref{fig-viability-supp}).

Third, we defined a version of the model with demographic and
environmental stochasticity, where environmental stochasticity was
represented by among-year variability in \(\sigma_{J_1}\). We used this
model to simulate a one-time introduction of 40 translocated adult
frogs. We ran this simulation 1000 times for each population and
computed the probability of a population becoming extinct after 50 years
given the observed parameter values and environmental stochasticity in
\(\sigma_{J_1}\). We varied the mean recruitment probability
\(\sigma_{J_1}\) from 0 and 0.25 and drew values of \(\sigma_{J_1}\)
each year from a beta distribution with a dispersion parameter of
\(\phi = 2\) (when \(\sigma_{J_1} = 0.5\) and \(\phi = 2\) the beta
distribution is uniform between 0 and 1). Using different values of
\(\phi\) does not qualitatively change the existence of distinct
extinction dynamics between populations with \(\sigma_{A_R} < 0.5\) and
those with \(\sigma_{A_R} > 0.5\). However, increasing yearly
variability in \(\sigma_{J_1}\) increases extinction risk for all
populations. For example, if we set \(\phi = 0.001\), such that in a
given year essentially either all year 1 juveniles survive or all of
them die, populations with \(\sigma_{A_R} > 0.5\) need to have
\(\sigma_{J_1}\) greater than 0.2 to have a 50-year extinction
probability of less 50\%. Because we do not pit tag juveniles, we do not
have CMR estimates for \(\sigma_{J_1}\) or \(\phi\). However, based on
qualitative and semi-quantitative field observations over 25 years, a
value of \(\sigma_{J_1} = 0.25\) in the presence of Bd is probably a
reasonable estimate for many populations. Thus, we expect our model
predictions to be conservative with regards to population recovery.

Finally, we assessed whether our stochastic model could reproduce
observed trajectories of population recovery. We focused on population
70550 because this was our longest CMR time series for a translocated
population and because this population shows evidence of substantial
post-translocation increases in adult abundance associated with
population establishment and recovery. We simulated our model for 16
years, repeating the simulation 50,000 times. For each run and each
year, we drew \(\sigma_{J_1}\) from a uniform distribution between 0 and
1 (or equivalently a beta distribution with mean 0.5 and \(\phi = 2\)).
Using Approximate Bayesian Computing and rejection sampling
\citep{kosmala2016}, we identified the top 2\% of trajectories (i.e.,
1000 trajectories) that minimized the sum of squared errors between the
observed and predicted data. The yearly \(\sigma_{J_1}\) values
associated with these ``best'' trajectories represented an approximate
posterior distribution \citep{beaumont2010}. Using these best fit
trajectories, we assessed whether our model could qualitatively describe
the patterns of recovery in the observed data for population 70550.

\hypertarget{frog-evolution-in-response-to-bd-2}{%
\section{Frog evolution in response to
Bd}\label{frog-evolution-in-response-to-bd-2}}

@@

\hypertarget{study-design}{%
\subparagraph{Study design}\label{study-design}}

We compared frog genomes sampled in naive versus recovering populations.
Comparing populations with different infection histories allowed larger
sample sizes and replication across the landscape. The alternative
approach of comparing samples from the same populations before and after
Bd exposure isn't feasible in this system because Bd arrived in most MYL
frog populations decades ago and population persistence/recovery is rare
and unpredictable. As a result, samples from recovering populations
collected from before and after Bd exposure are not available and are
unlikely to be available in the future.

\hypertarget{go-analysis}{%
\subparagraph{GO analysis}\label{go-analysis}}

In the Results section of the main paper, we describe the stringent set
of outlier SNPs (identified using a Bonferroni-corrected p-value of
0.01). The liberal set, identified using a Bonferroni-corrected p-value
of 0.05, included 38 outliers (35 SNPs and 3 INDELS) from 30 distinct
genes across 16 contigs. Using this liberal set of outlier variants, we
performed a statistical overrepresentation test (Fisher's exact test)
using Blast2GO \citep{conesa2005}. First, we retrieved the BLAST hits
and mapped GO terms for each gene in our targeted transcriptome. Next,
we compared the 30 unique outlier genes from the liberal set of 38
outlier variants to the complete set of genes in our target
transcriptome to determine if any GO biological functions, molecular
functions, or cellular processes were overrepresented. We then repeated
this process for the set of 33 genes located in the
\emph{F\textsubscript{ST}} and \(\pi_{diff}\) splined windows.

\hypertarget{genetic-diversity}{%
\subparagraph{Genetic diversity}\label{genetic-diversity}}

To characterize general patterns of genetic diversity between naive and
persistent populations, we calculated individual heterozygosity using
VCFtools \citep{danecek2011}, and compared average heterozygosity using
a Wilcoxon rank-sum test. Additionally, we used VCFtools to calculate
per-population \(\pi\) in 100kb sliding windows along the genome, and in
each of the 10 outlier windows identified in the splined window
analysis. We compared per-population average \(\pi\) values using
Wilcoxon rank-sum tests.

@@

\hypertarget{go-analysis-1}{%
\subparagraph{GO analysis}\label{go-analysis-1}}

We found no significantly over-represented GO terms in the liberal set
of outlier regions, or in the set of 33 genes located in the overlapping
\emph{F\textsubscript{ST}} and \(\pi_{diff}\) splined windows.
Therefore\ldots..

\hypertarget{genetic-diversity-1}{%
\subparagraph{Genetic diversity}\label{genetic-diversity-1}}

Average (i) individual-level heterozygosity and (ii) population-level
nucleotide diversity (\(\pi\)) were similar between naive and recovering
populations (Figure~\ref{fig-violinplot-heterozy};
Figure~\ref{fig-boxplot-pi}). The analysis of \(\pi\) across the 10
outlier windows indicates that within each of the windows there is
considerable variation in \(\pi\) between populations and between
population clusters (Figure~\ref{fig-barchart-pi-by-windowsite}).

\caption{\label{fig-yosemap}Map showing the locations of translocated
and donor MYL frog populations in Yosemite National Park (park boundary
indicated by gray polygon). Symbol shapes indicate the donor population
used for each translocation site, and 5-digit numbers identify each
donor and translocation site. To obscure the exact locations of
populations, random noise was added to all point coordinates. Inset map
shows the location of Yosemite within California. In both maps,
elevation is indicated by the colored hillshade layer (dark green =
lowest elevation, white = highest elevation).}

\caption{\label{fig-bdload-beforeafter}For frogs translocated to each of
the 12 recipient sites, Bd loads for the period immediately prior to
translocation versus during the 1-year period after translocation. Bd
loads are expressed as the number of ITS1 copies per skin swab, as
estimated by qPCR of the Bd ITS1 region. Box plots show medians, first
and third quartiles, largest and smallest values within 1.5x
interquartile range, and values outside the 1.5x interquartile range.
Loads indicative of severe disease are \textgreater{} 5.8 ITS copies (on
a log\textsubscript{10} scale). Samples sizes are provided immediately
above the x-axis.}

\caption{\label{fig-survival-postdens}Evidence that Bd load is not an
important predictor of post-translocation frog survival. Depicted
distributions are the estimated posterior density curves and shaded 95\%
uncertainty intervals for the intercept and all predictor variables from
the best model in the between-site meta-analysis. In the Bayesian
framework in which the model was developed, variables are considered
important predictors if the associated uncertainty interval does not
overlap zero (indicated by the dashed red line). Predictor variables
shown on the y-axis are defined as follows: length = frog size, snow\_t
= winter severity in the year of translocation (measured on April 1),
snow\_t1 = winter severity in the year following translocation (measured
on April 1), day = day of year on which a translocation was conducted,
bd\_load = Bd load, elevation = site elevation, order\_first =
within-site translocation order, sex\_male = frog sex, donor\_70567 and
donor\_72996 = donor population.}

\caption{\label{fig-compare_surv_probs}Comparison of average yearly
adult survival probabilities for adults translocated to each of 3 sites
versus adults that were naturally recruited at each site (as a result of
reproduction by translocated frogs). In contrast to
Fig. 3, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohorts. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. All estimates were derived using
the mrmr package, and the methods for calculation are described in
\textbf{Supporting Information - Population viability modeling --
Estimating model parameters}.}

\caption{\label{fig-viability-supp}Sensitivity analysis of the
stage-structured MYL frog model. Elasticity of \(\lambda\) with changes
in four parameters: \(\sigma_{A_R}\) (yearly survival probability of
naturally recruited adults), \emph{F (}number of eggs produced by a
female frog in a year that successfully hatch), \(\sigma_{J_1}\) (yearly
probability of survival of year 1 juveniles that also affects
recruitment), and \(\sigma_{J_2}\) (yearly probability of survival and
recruitment of year 2 juveniles). Elasticity is calculated at the
default parameter values for each population and \(\sigma_{J_1}\) =
0.09.}

@@

\caption{\label{fig-synteny-plot}Synteny plot showing conserved gene
order in \emph{Xenopus tropicalis} and \emph{Rana muscosa} for the
outlier region containing MHC Class I Classical and MHC Class III gene
regions. The plot was created with SimpleSynteny \citep{veltri2016}
using \emph{Xenopus tropicalis} Chromosome 8 (NC\_030684.2, genbank
accession GCA\_000004195.4) and \emph{Rana muscosa} Contig19. Asterices
indicate the location of SNP outliers in the TCF19 and VARS1 genes. Gap
sizes for each contig representation are labeled.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/het_by_status_violin_plot.png}

@@

\caption{\label{fig-violinplot-heterozy}Violin plots showing individual
heterozygosity for the Bd-naive and recovering populations included in
the frog genomics study. Individual data points are shown on the plot
with their corresponding site number (from
Fig. 1). A Wilcoxon signed-rank test showed
no significant difference between the two groups.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/boxplot_pi_by_site.png}

@@

\caption{\label{fig-boxplot-pi}For each study population in the frog
evolution study (from Fig. 1A), boxplots
showing nucleotide diversity (\(\pi\)) within 100kb sliding windows
along the genome. The y-axis has been log\textsubscript{10}-transformed
for display purposes. Sites are color-coded as ``Bd-naive'' or
``recovering'', and vertical dashed lines show the genetic breaks
between frog populations as described in
Figure~\ref{fig-selectionresults} A). Within the genetic clusters,
mean \(\pi\) is significantly different between sites (p \textless0.001
Wilcoxon rank-sum test).}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/perwindow_persite_pi.png}

@@

\caption{\label{fig-barchart-pi-by-windowsite}For each study population
in the frog evolution study, average \(\pi\) values by population
calculated for each of the 10 outlier windows identified in the splined
window analysis (A-J). Sites are color-coded as ``Bd-naive'' or
``recovering'', and vertical dashed lines show the genetic breaks
between frog populations as described in
Figure~\ref{fig-selectionresults} A. Gene annotations for the outlier
regions are as follows: (A) C6, C7, (B) DDX10, ZBTB24, (C) SULTR-like,
TRAF3IP2, VGLL3, EXOC1, (D) FLOT1, TUBB2B, TUBB, MDC1, CCHCR1,
TCF19\emph{, TRIM29, HSP70, LSM2, VARS,} (E) GCC2, CFAP251, PEG10, (F)
ERO1A, GVINP1, (G) ERO1A, GVINP1, (H) PPP1R12A, TSPAN4, PAWR, MFRP, MAX,
PPP6R3, (I) C6H5ORF22, PKS6, BSPRY, MPV17, and (J) CAD, ATRAID, GPN1.
Note that windows F and G are in the same gene region. * by gene names
indicate that an outlier SNP is also present in this gene . (K) Average
\(\pi\) values of naive and recovering groups for each of the outlier
windows shown in A-J. }
